
# Toy Example

That's define a cohort where:

* the index function returns the "start of the first $\heartsuit$ after the start of the first $\spadesuit$" (i.e $\text{index}(W) = \mathsf{F}^{\star}(W)$).

```{r}
index <- fstar
```

Based on the index feature, we want to extract the following features about $W$:

* 2spade1heart:
  * TRUE/FALSE: Anytime before index, does $W$ contain 2 $\spadesuit$ separated by at least 3 time units or 1 $\clubsuit$?
  * NOTE: this is analogous to a "2-out/1-in" algorithm
* foreverDiamonds:
  * TRUE/FALSE: Do the $\diamondsuit$ events have a gap of no more than 10 time units before index, starting from 25 units before index?
  * NOTE: This is analogous to a continuous enrollment algorithm
* flush: 
  * TRUE/FALSE: Is there a flush after index? I.e. any set of five consecutive events after index of the same suit.
  * Int: What is the time that the flush occurred?
  * Suit: What suit triggered the flush?
  * NOTE: the flush feature contains three data points: an indicator whether some event occurred, the time it occurred, and the type that triggered it.
* clubTrump:
  * TRUE/FALSE: Did a $\clubsuit$ occur after index?
  * Int: What is the time that the club occurred after index?
* fullOfHearts:
  * $\{Int\}$: the set of all times that $\heartsuit$ that occur before a clubTrump
  * NOTE: the feature in this case is a set of times (not just a single time)

Now let's express each of these features mathematically.


## 2spade1heart

Let's define a few more functions that operate on $w$ and $\Xi$.

| Definition     | Mapping                      | Description        |
| ------------   | ---------------------------- | ------------------ |
| $any(W) = \begin{cases} W \neq \emptyset & TRUE \\ else & FALSE \end{cases}$  | $any : \mathcal{W} \mapsto \mathbb{B}$ | Is $W$ empty? | 




## foreverDiamonds


$$
\begin{aligned}
\mathsf{L}^{\diamondsuit}(W) &= \{ w | w \in \mathsf{L}(W', \mathsf{P}^c(A_t, = \diamondsuit)  \} \\
\text{scangap} &= f(B) = \{(a', b') | (a, b) \in B, \,\, a' = \text{if }  x_{i} \leq y_{i - 1} + c \text{ then } x_{i - 1} \text{ else } x_{i} , b' = \text{if }  y_{i} \leq y_{i - 1} \text{ then } y_{i - 1} \text{ else } y_{i}, \forall i = 2, \dots, n  \} \\
\text{periods} &= g(B) = \{(a, b') | (a, b) \in f(B), \,\, b' = max(b) \,\, \forall a_i = a_j \} \\
\text{pdperiods} &= h(B, pd) = \{ (a, b) | (a, b) \in g(B), \,\, a > \mathsf{a}_s(pd), \, b < \mathsf{a}_e(pd) \} \\
\text{gaps} &= d(B, pd) = \{(a, b) | (a, b) \in B, TODO  \} \\
\text{foreverDiamonds}(W, \text{index}) &= \{\text{if } \{x | (a, b) \in d(h(A_{se}(\mathsf{L}^{\diamondsuit}), (-25, index)), x = b - a, x > 10 \} = \emptyset \text{ FALSE} \text { else } \text{TRUE} \} \\
\end{aligned}
$$

Let's look at the $\text{scangap}$ function and the patterns it can be composed of. First, we have a function performing the nuts and bolts of the set comprehension (e.g., $\text{if }  x_{i} \leq y_{i - 1} + c \text{ then } x_{i - 1} \text{ else } x_{i}$), which be somewhat kludglily patterned as:

```{r}
check_gap_cast <- function(cc = 0, pos = c(1, 2, 1, 1)) {
  function(a, b){
    if (b[pos[1]] <= (a[pos[2]] + cc)) a[pos[3]] else b[pos[4]]
  }
}
```

So that `check_gap_cast(cc = 10, pos = c(1, 2, 1, 1))` yields $f(x, y) = \text{if }  x_{i} \leq y_{i - 1} + 10 \text{ then } x_{i - 1} \text{ else } x_{i}$ and `check_gap_cast(cc = 0, pos = c(2, 2, 2, 2))` yields $f'(x, y) = \text{if }  y_{i} \leq y_{i - 1} \text{ then } y_{i - 1} \text{ else } y_{i}$.

Then we can apply that pattern to create the scangap function.

```{r}
scangap_cast <- function(cc){
  function(l){
    Reduce(
      f = function(x, y){
        c(check_gap_cast(cc, c(1, 2, 1, 1))(x, y), 
          check_gap_cast(0, c(2, 2, 2, 2))(x, y))
        },
    x = l, accumulate = TRUE)
  }
}

scangap_cast(10)(head(Wdt$xy))
```

In summary, scangap return a set of tuples the same length as the input $W$, where the start and end of each $w_i \in W$ have been updated depending on $w_{i - 1}$ extending the period until a time gap of 10 units occurs.

Now, we can identify which of those periods are the non overlapping periods by scanning the modified periods from the right, accumulating an indicator of whether each period is the longest segment of a period; then we can use set (list) comprehension to filter to the unique periods.

For example, (warning: this is klunky and could be cleaned up), but in `R`:

```{r}
# Finds the longest non-overlapping periods
scan_periods_cast <- function(cc){
  function(l){
     Map(
       function(x) as.logical(x[3]),
       Reduce(
         f = function(a, b) c(a, (a[2] + cc <= b[1]) * 1),
         x = l[-length(l)],
         init = c(l[[length(l)]], 1), accumulate = TRUE, right = TRUE)
     )
  }
}

period_cast <- function(cc){
  ff <- scan_periods_cast(cc)
  gg <- scangap_cast(cc)
  function(l){
    gg(l)[unlist(ff(gg(l)))]
  }
}

period_cast(10)(Wdt$xy)
period_cast(10)(head(Wdt$xy))
```

Now we have need a function that limits our periods to just those that overlap another period $pd$ (e.g. a lookback period). This is the $pdperiod$ function.

```{r}
pdperiod <- function(l, p){
  Filter(
    f = function(z) { !(p[2] < z[1]) && !(p[1] > z[2]) },
    x = l
  )  
}

period_cast(10)(Wdt$xy) %>% pdperiod(p = c(4, 10))
period_cast(10)(Wdt$xy) %>% pdperiod(p = c(-10, 10))
period_cast(10)(Wdt$xy) %>% pdperiod(p = c(-10, 90))
```

```{r, eval = FALSE}
# Define a couple of helper function
s <- function(p) p[1]
e <- function(p) p[2]

h <- function(p1, p2){
  if(s(p2) < s(p1) && e(p2) <= e(p1)){
    list(c(e(p2), e(p1)))
  } else if(s(p1) <= s(p2) && e(p1) < e(p2)){
    list(c(s(p1), s(p2)))
  } else if(s(p1) <= s(p2) && e(p2) <= e(p1)){
    list(c(s(p1), s(p2)), c(e(p2), e(p1)))
  } else {
    NULL
  }
}

f <- function(l, p2){
  out  <- l[-length(l)]
  p1   <- l[[length(l)]]
  hold <- h(p1, p2)
  append(out, hold)
}

h(c(1, 10), c(0, 5))
h(c(1, 10), c(2, 5))
h(c(1, 10), c(0, 11))

pds <- list(
  c(2, 5),
  c(10, 15),
  c(16, 17)
)

gap_cast <- function(pd){
  function(periods){
    Reduce(f, periods, init = list(pd))
  }
}


```


```{r, eval = FALSE}
shift_period <- function(p, x){
  x + p
}

lookback <- shift_period(c(-25, 0), index(W))

preindex_gaps <- gap_cast(lookback)

# period_cast(10)(Wdt$xy) %>% 
#   clip_period(p = lookback) 
# 
# h(lookback, c(1, 67))
# %>%
#   Reduce(f, ., init = list(lookback))
```


Now let's build foreverDiamonds
```{r, eval = FALSE}
Ldiamond <- Lsuit_cast("diamond")

foreverDiamonds <- function(W){
  Ldiamond(W) %>% 
    A_prd %>% 
    { period_cast(10)(.) } %>%
    preindex_gaps %>%
    { length(.) > 0}
}

foreverDiamonds(W)
```

Note that another pattern emerged! `Lw` is exactly the same pattern as the first part of `index`! 

## fullHouse

* $\text{fullHouse}(W, \text{index})$

## postClubs

* $\text{postClubs}(W, \text{index})$

